\chapter{Logic and Propositions as Types}

In this chapter, we develop the logical and type-theoretic foundations
that underpin Lean.
We begin with propositional logic and its formalization via natural
deduction (Section~\ref{sec:prop_logic}).
We then show how types are constructed using analogous inference rules
(Section~\ref{sec:primitive_types}).
In Section~\ref{sec:curry_howard}, we make the connection explicit
via the Curry--Howard correspondence.
Section~\ref{sec:predicate_logic} extends the picture to predicate
logic and dependent types, and Section~\ref{sec:constructive} discusses
the distinction between constructive and classical reasoning.

\section{Propositional Logic and Natural Deduction}\label{sec:prop_logic}

Logic is the study of reasoning, branching into various systems.
We refer to \textbf{classical logic} as the logical system that
underpins much of traditional mathematics.
We first introduce \textbf{propositional logic}, the simplest form of
classical logic.
Later, in Section~\ref{sec:predicate_logic}, we extend this to
\textbf{predicate (or first-order) logic}, which adds
\textbf{predicates} and \textbf{quantifiers}.

In propositional logic, a \textbf{proposition} is a statement that is
either true or false, and a \textbf{proof} is a logical argument that
establishes the truth of a proposition.
Propositions can be combined with logical \textbf{connectives} such as
``and'' ($\wedge$),
``or'' ($\vee$),
``not'' ($\neg$),
``false'' ($\bot$),
``true'' ($\top$),
``implies'' ($\Rightarrow$),
and ``if and only if'' ($\Leftrightarrow$).
These connectives allow the creation of compound propositions.
Here is how some connectives are represented in Lean:

\begin{example}[Logical connectives in Lean]\mbox{}
  \begin{lstlisting}[language=lean]
    #check @And  -- And : Prop → Prop → Prop
    #check @Or   -- Or  : Prop → Prop → Prop
  \end{lstlisting}
  \lstinline[language=lean]|Prop| is the type of propositions,
  introduced in Chapter~1.
\end{example}

Logic is often formalized through a framework known as the
\textbf{natural deduction system}, developed by Gentzen in the 1930s
(\cite{wadler2015propositions}).
This approach specifies rules for deriving \textbf{conclusions} from
\textbf{premises}, called \textbf{inference rules}.

\begin{example}[Inference rule (schematic)]
  Here is a schematic inference rule
  (\cite{nordstrom1990programming}, page~35):
  \begin{prooftree}
    \AxiomC{$P_1$}
    \AxiomC{$P_2$}
    \AxiomC{$\cdots$}
    \AxiomC{$P_n$}
    \QuaternaryInfC{$C$}
  \end{prooftree}
  The propositions $P_1, P_2, \ldots, P_n$ above the line are the
  premises, and the conclusion $C$ appears below the line.
\end{example}

The inference rules we need are \textbf{introduction rules}, which
specify how to form compound propositions from simpler ones, and
\textbf{elimination rules}, which allow us to extract information
from compound propositions.
We now present the rules for the main connectives, following
\cite{thompson1999types}, Section~1.1.

\subsection*{Conjunction ($\land$)}

\paragraph{Introduction Rule.}
If we have proofs of both $A$ and $B$, we can form a proof of
$A \land B$:
\begin{prooftree}
  \AxiomC{$A$}
  \AxiomC{$B$}
  \RightLabel{$\land$-Intro}
  \BinaryInfC{$A \land B$}
\end{prooftree}

\paragraph{Elimination Rules.}
From a proof of $A \land B$, we can extract a proof of either
conjunct:

\noindent
\begin{minipage}[t]{0.48\textwidth}
  \vspace{0pt}
  \begin{prooftree}
    \AxiomC{$A \land B$}
    \RightLabel{$\land$-Elim$_1$}
    \UnaryInfC{$A$}
  \end{prooftree}
\end{minipage}\hfill
\begin{minipage}[t]{0.48\textwidth}
  \vspace{0pt}
  \begin{prooftree}
    \AxiomC{$A \land B$}
    \RightLabel{$\land$-Elim$_2$}
    \UnaryInfC{$B$}
  \end{prooftree}
\end{minipage}

\subsection*{Disjunction ($\lor$)}

\paragraph{Introduction Rules.}
A proof of either $A$ or $B$ suffices to form a proof of
$A \lor B$:

\noindent
\begin{minipage}[t]{0.48\textwidth}
  \vspace{0pt}
  \begin{prooftree}
    \AxiomC{$A$}
    \RightLabel{$\lor$-Intro$_1$}
    \UnaryInfC{$A \lor B$}
  \end{prooftree}
\end{minipage}\hfill
\begin{minipage}[t]{0.48\textwidth}
  \vspace{0pt}
  \begin{prooftree}
    \AxiomC{$B$}
    \RightLabel{$\lor$-Intro$_2$}
    \UnaryInfC{$A \lor B$}
  \end{prooftree}
\end{minipage}

\paragraph{Elimination Rule (proof by cases).}
If $C$ follows from $A$ and also from $B$, then $C$ follows from
$A \lor B$:
\begin{prooftree}
  \AxiomC{$A \lor B$}
  \AxiomC{$[A] \vdash C$}
  \AxiomC{$[B] \vdash C$}
  \RightLabel{$\lor$-Elim}
  \TrinaryInfC{$C$}
\end{prooftree}

\subsection*{Implication ($\to$)}

\paragraph{Introduction Rule.}
To prove $A \to B$, we assume $A$ and derive $B$:
\begin{prooftree}
  \AxiomC{$[A] \vdash B$}
  \RightLabel{$\to$-Intro}
  \UnaryInfC{$A \to B$}
\end{prooftree}

\paragraph{Elimination Rule (modus ponens).}
Given a proof of $A \to B$ and a proof of $A$, we obtain a proof
of $B$:
\begin{prooftree}
  \AxiomC{$A \to B$}
  \AxiomC{$A$}
  \RightLabel{$\to$-Elim}
  \BinaryInfC{$B$}
\end{prooftree}

\begin{notation}
  We use $A \vdash B$ (called the \textbf{turnstile}) to denote a
  derivation of $B$ from $A$; it can be read as ``$A$ entails $B$.''
  The square brackets around a premise, as in $[A]$, indicate that
  the assumption $A$ is \textbf{discharged} (i.e., released) at the
  conclusion of the rule.
  To illustrate: the $\to$-Intro rule says that to prove
  $A \to B$, we temporarily assume $A$, derive $B$ under this
  assumption, and then \emph{discharge} the assumption---meaning
  that the conclusion $A \to B$ no longer depends on $A$ being true.
  The assumption was only needed during the derivation and is consumed
  by the rule.
\end{notation}

Having introduced the rules of propositional logic, we now turn to
the type-theoretic side, where analogous rules govern the
construction of types.

\section{Primitive Types}\label{sec:primitive_types}

Type theory employs a similar system of inference rules, but
expressed in terms of \textbf{type judgments}.
A type judgment has the form $\Gamma \vdash t : T$,
meaning: under \textbf{context} $\Gamma$ (a list of typed variables),
the term $t$ has type $T$.
Introduction and elimination rules are then inference rules that
derive new type judgments from existing ones, allowing us to construct
new compound types from existing ones.

\begin{example}[Type judgment rule (schematic)]
  A schematic type judgment rule has the following form:
  \begin{prooftree}
    \AxiomC{$\Gamma \vdash a : A$}
    \AxiomC{$\Gamma \vdash b : B$}
    \BinaryInfC{$\Gamma \vdash c : C$}
  \end{prooftree}
  Reading: if, under context $\Gamma$, we have a term $a$ of type $A$
  and a term $b$ of type $B$, then we can construct a term $c$ of
  type $C$.
\end{example}

Technically, there are two additional kinds of rules that we will not
treat in detail here:
\textbf{formation rules}, which declare that a type is well-defined,
and \textbf{computation rules}, which specify how terms are evaluated.

Moreover, there is a specific form of judgment for equality:
$\Gamma \vdash A \equiv B\ \text{type}$ means ``types $A$ and $B$ are
\textbf{judgmentally (or definitionally) equal} in context $\Gamma$.''
Similarly for terms: $\Gamma \vdash t_1 \equiv t_2 : A$ means
``terms $t_1$ and $t_2$ are judgmentally equal of type $A$ in
context $\Gamma$.''

\paragraph{Brief explanation of equality in type theory.}

It is important to distinguish two notions of equality in type theory.

\textbf{Definitional (or judgmental) equality} is a relation that
Lean's kernel can verify automatically, without any additional proof
steps.
Informally, two terms are definitionally equal when the kernel can
reduce them to the same normal form by applying computation rules and
unfolding definitions.
A \textbf{normal form} is the most reduced state of a term, obtained
by systematically applying reduction rules until no further reductions
are possible.
For example, $2 + 2$ and $4$ are definitionally equal because both
reduce to the same normal form.

In Lean, the operator \lstinline[language=lean]|:=| is a
\textbf{syntactic assignment}: it creates a new definition by binding
a name to a term.
For instance, writing \lstinline[language=lean]|def x : Nat := 2 + 2|
introduces a new name \lstinline[language=lean]|x| and makes
\lstinline[language=lean]|x| definitionally equal to
\lstinline[language=lean]|2 + 2| (and hence to
\lstinline[language=lean]|4|).
Note that \lstinline[language=lean]|:=| is not a predicate that can be
evaluated as true or false; it is part of the syntax for creating
definitions.

In contrast, \textbf{propositional equality} is a statement
\emph{within} the logic (expressed as a type, e.g.,
\lstinline[language=lean]|a = b|) that may require an explicit proof.
Because it is a logical proposition rather than a computation,
propositional equality is not directly decidable by the type checker
and must be established by the user.

Let us now construct new types from given types $A$ and $B$.

\subsection*{Product Type}

The type $A \times B$ denotes the type of pairs $(a, b)$ where
$a : A$ and $b : B$, called the \textbf{product type}.

\paragraph{Introduction Rule (pairing).}
\begin{prooftree}
  \AxiomC{$a : A$}
  \AxiomC{$b : B$}
  \BinaryInfC{$(a, b) : A \times B$}
\end{prooftree}

\noindent In Lean:
\begin{lstlisting}[language=lean]
Prod.mk a b : Prod A B   -- or A × B
(a, b) : A × B
⟨a, b⟩ : A × B
\end{lstlisting}

\paragraph{Elimination Rules (projections).}\mbox{}\\[0.5em]
\noindent
\begin{minipage}[t]{0.48\textwidth}
  \vspace{0pt}
  \begin{prooftree}
    \AxiomC{$p : A \times B$}
    \UnaryInfC{$\mathsf{fst}(p) : A$}
  \end{prooftree}
\end{minipage}\hfill
\begin{minipage}[t]{0.48\textwidth}
  \vspace{0pt}
  \begin{prooftree}
    \AxiomC{$p : A \times B$}
    \UnaryInfC{$\mathsf{snd}(p) : B$}
  \end{prooftree}
\end{minipage}

\noindent In Lean:
\begin{lstlisting}[language=lean]
  p.1 : A       -- or Prod.fst p
  p.2 : B       -- or Prod.snd p
\end{lstlisting}

\subsection*{Sum Type}

The \textbf{sum type} $A + B$ (also called a coproduct or disjoint
union) consists of values that are either of type $A$ (tagged with
$\mathsf{inl}$) or of type $B$ (tagged with $\mathsf{inr}$).

\paragraph{Introduction Rules (injections).}\mbox{}\\[0.5em]
\noindent
\begin{minipage}[t]{0.48\textwidth}
  \vspace{0pt}
  \begin{prooftree}
    \AxiomC{$a : A$}
    \UnaryInfC{$\mathsf{inl}(a) : A + B$}
  \end{prooftree}
\end{minipage}\hfill
\begin{minipage}[t]{0.48\textwidth}
  \vspace{0pt}
  \begin{prooftree}
    \AxiomC{$b : B$}
    \UnaryInfC{$\mathsf{inr}(b) : A + B$}
  \end{prooftree}
\end{minipage}

\noindent In Lean:
\begin{lstlisting}[language=lean]
Sum.inl a : Sum A B   -- or A ⊕ B
Sum.inr b : Sum A B
\end{lstlisting}

\paragraph{Elimination Rule (case analysis).}
\begin{prooftree}
  \AxiomC{$p : A + B$}
  \AxiomC{$f : A \to C$}
  \AxiomC{$g : B \to C$}
  \TrinaryInfC{$\mathsf{cases}(p, f, g) : C$}
\end{prooftree}

\noindent In Lean:
\begin{lstlisting}[language=lean]
example (p : Sum A B) (f : A → C) (g : B → C) : C := by
  cases p with
  | inl x => exact f x
  | inr y => exact g y
\end{lstlisting}

\subsection*{Function Types}

The type $A \to B$ represents functions from $A$ to $B$.

\paragraph{Introduction Rule (lambda abstraction).}
\begin{prooftree}
  \AxiomC{$x : A \vdash \Phi : B$}
  \UnaryInfC{$\lambda x.\,\Phi : A \to B$}
\end{prooftree}

\noindent
Here $\lambda x.\,\Phi$ is a function that maps any element $x : A$
to the element $\Phi : B$.
In Lean, lambda abstraction is written using
\lstinline[language=lean]|fun| or \lstinline[language=lean]|λ|:
\begin{lstlisting}[language=lean]
def identityFun (A : Type) : A → A := fun x => x
\end{lstlisting}

\paragraph{Elimination Rule (function application).}
\begin{prooftree}
  \AxiomC{$f : A \to B$}
  \AxiomC{$a : A$}
  \BinaryInfC{$f(a) : B$}
\end{prooftree}

\noindent In Lean, function application is written by juxtaposition:
\begin{lstlisting}[language=lean]
example (f : A → B) (a : A) : B := f a
\end{lstlisting}

Functions are a primitive concept in type theory.
We can \textbf{apply} a function $f : A \to B$ to an element $a : A$
to obtain an element of $B$, denoted $f(a)$.
In type theory, it is common to omit the parentheses and write the
application simply as $f\, a$.

Having introduced the basic type constructions, we now make explicit
the correspondence between these types and the logical connectives
from Section~\ref{sec:prop_logic}.

\section{The Curry--Howard Correspondence}\label{sec:curry_howard}

The reader will have noticed a strong similarity between the
inference rules for logical connectives
(Section~\ref{sec:prop_logic}) and the rules for constructing types
(Section~\ref{sec:primitive_types}).
For instance, function types mirror implications, product types
mirror conjunctions, and sum types mirror disjunctions.
This is not a coincidence: it is an instance of the
\textbf{Curry--Howard correspondence}, a deep connection between
logic and computation discovered independently by Haskell Curry
(1930s--50s) and William Howard (1969).
It establishes a systematic dictionary between proofs and programs (\cite{thompson1999types}):

\begin{center}
  \begin{tabular}{lll}
    \textbf{Logic}                &                       & \textbf{Type Theory}      \\
    \hline
    Implication $P \Rightarrow Q$ & $\longleftrightarrow$ & Function type $P \to Q$   \\
    Conjunction $P \land Q$       & $\longleftrightarrow$ & Product type $P \times Q$ \\
    Disjunction $P \lor Q$        & $\longleftrightarrow$ & Sum type $P + Q$          \\
    Proposition                   & $\longleftrightarrow$ & Type                      \\
    Proof                         & $\longleftrightarrow$ & Term                      \\
  \end{tabular}
\end{center}

\noindent
A proof of an \textbf{implication} $P \Rightarrow Q$ is a function
that transforms any proof of $P$ into a proof of $Q$.
A proof of a \textbf{conjunction} $P \land Q$ consists of a pair
containing proofs of both conjuncts.
A proof of a \textbf{disjunction} $P \lor Q$ is either a proof of
$P$ or a proof of $Q$.

Lean realizes this correspondence concretely.
For instance, $A \land B$ can be represented as
\lstinline[language=lean]|And A B| or \lstinline[language=lean]|A ∧ B|.
Its introduction rule is
\lstinline[language=lean]|And.intro _ _| or simply
\lstinline[language=lean]|⟨_, _⟩| (underscores are placeholders for
the two proofs).
The compound proposition $A \land B$ can then be decomposed using the
elimination rules
\lstinline[language=lean]|And.left| and
\lstinline[language=lean]|And.right|.

\begin{example}\label{ex:conj_intro_2}
  Let us look at a simple Lean example:
  \begin{lstlisting}[language=lean]
    example {a b : Prop} (ha : a) (hb : b) : (a ∧ b) := And.intro ha hb
  \end{lstlisting}
  Using curly brackets \lstinline[language=lean]|{ }|, we let Lean
  infer that $a$ and $b$ are propositions
  (\lstinline[language=lean]|Prop|).
  The example states that given a proof of $a$
  (\lstinline[language=lean]|ha|) and a proof of $b$
  (\lstinline[language=lean]|hb|), we can form a proof of
  $a \land b$.
  The type of \lstinline[language=lean]|And.intro| is:
  \begin{lstlisting}[language=lean]
    And.intro : a → b → a ∧ b
  \end{lstlisting}
  It says: given a proof of $a$ and a proof of $b$, it returns a
  proof of $a \land b$.
  Here is an equivalent way to write the same statement using
  anonymous constructor notation:
  \begin{lstlisting}[language=lean]
    example {a b : Prop} (ha : a) (hb : b) : a ∧ b := ⟨ha, hb⟩
  \end{lstlisting}
\end{example}

\noindent
For a more concrete example, let us see how building a proof tree in
natural deduction corresponds to constructing a term in Lean.
To prove a compound statement, one typically follows a
\textbf{top-down} approach, decomposing the goal step by step.

\begin{example}[Associativity of Conjunction]
  We prove that $(A \land B) \land C$ implies $A \land (B \land C)$.
  First, from the assumption $(A \land B) \land C$, we can derive $A$:
  \begin{prooftree}
    \AxiomC{$(A \land B) \land C$}
    \RightLabel{$\land E_1$}
    \UnaryInfC{$A \land B$}
    \RightLabel{$\land E_1$}
    \UnaryInfC{$A$}
  \end{prooftree}
  Second, we can derive $B \land C$:
  \begin{prooftree}
    \AxiomC{$(A \land B) \land C$}
    \RightLabel{$\land E_1$}
    \UnaryInfC{$A \land B$}
    \RightLabel{$\land E_2$}
    \UnaryInfC{$B$}
    \AxiomC{$(A \land B) \land C$}
    \RightLabel{$\land E_2$}
    \UnaryInfC{$C$}
    \RightLabel{$\land I$}
    \BinaryInfC{$B \land C$}
  \end{prooftree}
  Finally, combining these derivations we obtain $A \land (B \land C)$:
  \begin{prooftree}
    \AxiomC{$(A \land B) \land C \vdash A$}
    \AxiomC{$(A \land B) \land C \vdash B \land C$}
    \RightLabel{$\land I$}
    \BinaryInfC{$A \land (B \land C)$}
  \end{prooftree}
\end{example}

\begin{example}[Lean Implementation]
  Let us now implement the same proof in Lean.
  \begin{lstlisting}[language=lean]
theorem and_associative {a b c : Prop} : (a ∧ b) ∧ c → a ∧ (b ∧ c) :=
  fun h : (a ∧ b) ∧ c ↦
  -- First, from the assumption (a ∧ b) ∧ c, we can derive a:
  have hab : a ∧ b := h.left
  have ha : a := hab.left
  -- Second, extract b and c:
  have hc : c := h.right
  have hb : b := hab.right
  -- Finally, combining these derivations we obtain a ∧ (b ∧ c):
  show a ∧ (b ∧ c) from ⟨ha, ⟨hb, hc⟩⟩
  \end{lstlisting}
  We introduce the \lstinline[language=lean]|theorem| with the name
  \lstinline[language=lean]|and_associative|.
  The type signature
  \lstinline[language=lean]|(a ∧ b) ∧ c → a ∧ (b ∧ c)| represents
  our logical implication.
  We construct the proof of the implication as a function (following
  the Curry--Howard correspondence) using the
  \lstinline[language=lean]|fun| keyword.
  The \lstinline[language=lean]|have| keyword introduces local lemmas
  within our proof, allowing us to break down complex reasoning into
  manageable intermediate steps (mirroring the natural deduction
  proof tree from before).

  Just before the \lstinline[language=lean]|show| keyword, the
  infoview displays the following context and goal:
  \begin{lstlisting}[language=lean]
  a b c : Prop
  h : (a ∧ b) ∧ c
  hab : a ∧ b
  ha : a
  hc : c
  hb : b
  ⊢ a ∧ b ∧ c
  \end{lstlisting}
  Resembling type judgments, the goal appears after the turnstile
  ($\vdash$), and the hypotheses above it form the current context.

  Finally,
  \lstinline[language=lean]|show a ∧ (b ∧ c) from ⟨ha, ⟨hb, hc⟩⟩|
  asserts that we are constructing a proof of
  \lstinline[language=lean]|a ∧ (b ∧ c)| using the term
  \lstinline[language=lean]|⟨ha, ⟨hb, hc⟩⟩|.
  The \lstinline[language=lean]|show| keyword improves readability and
  checks that the stated type matches the current goal.
  Here, the goal displayed as
  \lstinline[language=lean]|⊢ a ∧ b ∧ c| and the statement
  \lstinline[language=lean]|a ∧ (b ∧ c)| are \emph{syntactically
    identical}, because \lstinline[language=lean]|∧| is
  right-associative: no reduction rules are needed.
  If we had used \lstinline[language=lean]|show| with a type that was
  only \emph{propositionally} equal to the goal (but not
  definitionally equal), Lean would reject it.
\end{example}

We have now seen the Curry--Howard correspondence for propositional
connectives.
To capture more complex mathematical ideas, we need to extend the
picture to predicates and quantifiers.

\section{Predicate Logic and Dependent Types}\label{sec:predicate_logic}

To capture more complex mathematical ideas, we extend our system from
propositional logic to \textbf{predicate logic}.
A \textbf{predicate} is a proposition that depends on a variable.
In propositional logic we represent a proposition simply by $P$.
In predicate logic, a predicate is written as $P(a)$,
where $a$ is a variable.
Notice that a predicate is simply a
function from elements to propositions.

This extension allows us to introduce \textbf{quantifiers}:
$\forall$ (``for all'') and $\exists$ (``there exists'').
These quantifiers express that a given property holds either for
every element or for at least one element, respectively.
In Lean, if \lstinline[language=lean]|α| is any type, we can
represent a predicate \lstinline[language=lean]|P| on
\lstinline[language=lean]|α| as an object of type
\lstinline[language=lean]|α → Prop|.
Thus, given \lstinline[language=lean]|x : α|, the expression
\lstinline[language=lean]|P x : Prop| represents the proposition
that $P$ holds for $x$.

We can give an informal reading of the quantifiers as generalized
logical operations:
\begin{align*}
  \forall x.\,P(x) & \;\sim\; P(a) \land P(b) \land P(c) \land \ldots \\
  \exists x.\,P(x) & \;\sim\; P(a) \lor P(b) \lor P(c) \lor \ldots
\end{align*}
The dot symbol following the quantifier, as in $\forall x.\,$, binds
every occurrence of the variable $x$ in the expression $P(x)$.
The expression $\forall x.\, P(x)$ can be understood as a
generalized conjunction: $P$ holds for all possible values of $x$.
Similarly, $\exists x.\, P(x)$ is a generalized disjunction: $P$
holds for at least one value of $x$.

Under the Curry--Howard correspondence, universal quantification
corresponds to \textbf{dependent function types} (also called Pi
types, written $\Pi$), while existential quantification corresponds
to \textbf{dependent pair types} (also called Sigma types, written
$\Sigma$).
These are constructs from dependent type theory, in which types may
depend on values.

In Lean, \lstinline[language=lean]|∀| is the built-in dependent
function type (Pi type), while \lstinline[language=lean]|∃| is
defined as an inductive type.
We now present the Lean syntax for working with quantifiers directly.

\begin{example}[Quantifiers in Lean]
  Lean expresses quantifiers as follows:
  \begin{lstlisting}[language=lean]
variable (X : Type) (P : X → Prop)
#check (∀ (x : X), P x)  -- Prop (corresponds to Pi type Π)
#check (∃ (x : X), P x)  -- Prop (corresponds to Sigma type Σ)
  \end{lstlisting}
\end{example}

\begin{example}[Universal introduction in Lean]
  The \textbf{universal introduction rule} allows us to prove
  $\forall x,\, P(x)$ by proving $P(x)$ for an \textbf{arbitrary}
  $x$.
  In Lean, this corresponds to constructing a dependent function:
  \begin{lstlisting}[language=lean]
  example : ∀ n : Nat, n ≥ 0 :=
    fun n => Nat.zero_le n
  \end{lstlisting}
  Here \lstinline[language=lean]|Nat.zero_le| is a built-in lemma
  proving that $0 \le n$ for every natural number $n$.
\end{example}

\begin{example}[Universal elimination in Lean]
  The \textbf{universal elimination rule} allows us to instantiate a
  universally quantified statement with a specific value.
  In Lean, this is simply function application:
  \begin{lstlisting}[language=lean]
  example (h : ∀ n : Nat, n ≥ 0) : 5 ≥ 0 :=
    h 5
  \end{lstlisting}
\end{example}
\newpage
\begin{example}[Existential introduction in Lean]
  To prove an existential statement, we provide a \textbf{witness}
  together with a proof that the witness satisfies the predicate:
  \begin{lstlisting}[language=lean]
  example (x : Nat) (h : x > 0) : ∃ y, y < x :=
    ⟨0, h⟩
  \end{lstlisting}
  The term \lstinline[language=lean]|⟨0, h⟩| is a dependent pair
  holding the witness ($0$) and a proof that $0 < x$.
\end{example}

\begin{example}[Existential elimination in Lean]
  The \textbf{existential elimination rule} allows us to use an
  existential hypothesis $\exists x,\, P(x)$ by extracting a witness
  and the associated proof.
  In Lean, this can be done in term mode using pattern matching
  via \lstinline[language=lean]|match|:
  \begin{lstlisting}[language=lean]
  example (h : ∃ n : Nat, n + n = 10) : ∃ m : Nat, 10 = m + m :=
    match h with
    | ⟨n, hn⟩ => ⟨n, hn.symm⟩
  \end{lstlisting}
  The \lstinline[language=lean]|match| expression destructs the
  existential hypothesis \lstinline[language=lean]|h| into a witness
  \lstinline[language=lean]|n : Nat| and a proof
  \lstinline[language=lean]|hn : n + n = 10|.
  We then construct the new existential proof with the same witness
  \lstinline[language=lean]|n| and the reversed equality
  \lstinline[language=lean]|hn.symm : 10 = n + n|.
  This mirrors the logical pattern: given
  $\exists n,\, n + n = 10$, we produce $\exists m,\, 10 = m + m$
  using the same witness.
\end{example}

We have now extended the Curry--Howard correspondence from
propositional connectives to quantifiers.
In the next section, we discuss an important foundational
distinction: the difference between constructive and classical
reasoning, and how Lean navigates between the two.

\section{Constructive and Classical Reasoning}\label{sec:constructive}

A common pairing in the foundations of mathematics is:
\emph{set theory with classical logic} on one side, and
\emph{type theory with constructive logic} on the other.
While this pairing is the most common in practice, it is important to
note that neither association is inherent.
Intuitionistic set theories (such as IZF) exist, and on the other
hand, Russell and Whitehead's type theory in
\emph{Principia Mathematica} (historically the first type
theory) is entirely classical.
Lean's type theory is constructive by default, but classical axioms
can be invoked when needed.

A key point of departure between classical and constructive logic
concerns the \textbf{law of excluded middle} (LEM): the principle
that for every proposition $P$, either $P$ or $\lnot P$ holds.
In classical logic, LEM is taken as an axiom, so every proposition is
either true or false regardless of whether we can prove which.
In constructive logic, asserting $P \lor \lnot P$ requires actually
producing either a proof of $P$ or a proof of $\lnot P$.
For an open conjecture such as the Riemann Hypothesis, where neither
a proof nor a disproof is currently known, the constructive
mathematician cannot assert this disjunction.

Constructive logic emphasizes that a statement is considered true
only if we can explicitly provide a \textbf{witness} or
\textbf{construction} for it.
This is what makes constructive mathematics inherently
\textbf{computable}: a proof of $\exists x,\, P(x)$ must include a
concrete $x$ together with a proof that $P(x)$ holds.

\paragraph{Proof by contradiction.}
It is important to distinguish two forms of reasoning that are often
both called ``proof by contradiction'':
\begin{enumerate}
  \item \textbf{Proof of negation} (constructively valid): to prove
        $\lnot P$, assume $P$ and derive a contradiction ($\bot$).
        This is simply the definition of $\lnot P$ as $P \to \bot$ and
        is fully accepted in constructive logic.
  \item \textbf{Classical contradiction}: to prove $P$, assume
        $\lnot P$ and derive a contradiction, then conclude $P$.
        This requires LEM and is \emph{not} constructively valid.
\end{enumerate}

\begin{example}[Constructive existence proof]
  We give a constructive proof in Lean that there exist natural
  numbers $a$ and $b$ such that $a + b = 7$:
  \begin{lstlisting}[language=lean]
example : ∃ a b : Nat, a + b = 7 := ⟨3, 4, rfl⟩
  \end{lstlisting}
  To prove the existential statement, we provide witnesses
  ($a = 3$ and $b = 4$) and a proof that the predicate holds
  ($3 + 4 = 7$, verified by \lstinline[language=lean]|rfl|).
\end{example}

\paragraph{Ex falso quodlibet.}
The principle \textbf{ex falso quodlibet} (from a contradiction,
anything follows) is \emph{fully valid in constructive logic}.
It corresponds to the elimination rule for the empty type
$\bot$: if we have a proof of $\bot$ (falsehood), we can derive any
proposition.
In Lean, the \lstinline[language=lean]|exfalso| tactic applies this
principle:

\begin{example}[Reasoning from false]
  Here is an example of deriving any proposition from a contradiction:
  \begin{lstlisting}[language=lean]
  example (p : Prop) (h : False) : p := by
    exfalso
    exact h
  \end{lstlisting}
  The \lstinline[language=lean]|exfalso| tactic transforms the goal
  into $\vdash \mathsf{False}$, meaning we must now produce a proof
  of falsehood.
  Since we already have $h : \mathsf{False}$ as a hypothesis, we
  conclude with \lstinline[language=lean]|exact h|.
\end{example}

\paragraph{Classical reasoning in Lean.}
When genuinely classical reasoning is required, Lean provides the
tactic \lstinline[language=lean]|by_contra|, which invokes the law of
excluded middle (available in Lean as
\lstinline[language=lean]|Classical.em|).
This tactic allows us to assume $\lnot P$ and derive a
contradiction in order to conclude $P$:

\begin{example}[Classical proof by contradiction]
  \begin{lstlisting}[language=lean]
  example (p : Prop) (h : ¬¬p) : p := by
    by_contra hnp
    exact h hnp
  \end{lstlisting}
  The tactic \lstinline[language=lean]|by_contra hnp| introduces the
  hypothesis \lstinline[language=lean]|hnp : ¬p| and changes the goal
  to $\vdash \mathsf{False}$.
  We then apply the double negation hypothesis $h$ to
  \lstinline[language=lean]|hnp| to obtain the required
  contradiction.
  This proof is not constructively valid since it relies on the law of
  excluded middle.
\end{example}

In this chapter, we have seen how logical connectives and quantifiers
correspond to type-theoretic constructions via the Curry-Howard
correspondence, and how Lean blends constructive and classical
reasoning.
In the next chapter, we turn to how mathematical objects---from
number systems to algebraic structures---are modeled in Lean and
Mathlib.