\documentclass[12pt,a4paper]{book}
\usepackage{thesis}
\usepackage{hyperref}
\usepackage{stmaryrd} 
\usepackage{listings}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{bussproofs}
\usepackage{color}
\usepackage[backend=biber,style=alphabetic]{biblatex}
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}


\addbibresource{bibliography.bib}

\definecolor{keywordcolor}{rgb}{0.7, 0.1, 0.1}   % red
\definecolor{tacticcolor}{rgb}{0.0, 0.1, 0.6}    % blue
\definecolor{commentcolor}{rgb}{0.4, 0.4, 0.4}   % grey
\definecolor{symbolcolor}{rgb}{0.0, 0.1, 0.6}    % blue
\definecolor{sortcolor}{rgb}{0.1, 0.5, 0.1}      % green
\definecolor{attributecolor}{rgb}{0.7, 0.1, 0.1} % red
\def\lstlanguagefiles{lstlean.tex}
% set default language
\lstset{language=lean}

% Optional packages for your specific needs
% \usepackage{tikz}           % For diagrams
% \usepackage{pgfplots}       % For plots
% \usepackage{listings}       % For code listings
% \usepackage{algorithm2e}    % For algorithms
% \usepackage{hyperref}       % For hyperlinks (load last)

% Thesis metadata - CUSTOMIZE THESE
\title{Introduction to Formal Mathematics in Lean  with an example in Topology}
\author{Daniele Bolla}
\matrikelnr{21-694-187}
\supervisor{Prof. David Loeffler}
\submitdate{December 31, 2025}
% \thesistype{Bachelor Thesis}  % Optional: defaults to "Bachelor Thesis"
% \faculty{Faculty of Mathematics and Computer Science}  % Optional: already set
% \universitylogo{FernUni}      % Optional: path to logo file

% Theorem environments
% \newtheorem{thm}{Theorem}[chapter]
% \newtheorem{cor}[thm]{Corollary}
% \newtheorem{prop}[thm]{Proposition}
% \newtheorem{lem}[thm]{Lemma}
% \theoremstyle{definition}
% \newtheorem{defn}[thm]{Definition}
% \newtheorem{axiom}[thm]{Axiom}
% \theoremstyle{remark}
% \newtheorem{rmk}[thm]{Remark}
% \newtheorem{exmp}[thm]{Example}
% \newtheorem{exer}[thm]{Exercise}

% Custom commands for mathematics
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\set}[1]{\{#1\}}
\newcommand{\abs}[1]{\lvert#1\rvert}
\newcommand{\norm}[1]{\lVert#1\rVert}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}  
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{note}[theorem]{Note}
\numberwithin{figure}{theorem}
\begin{document}

% Title page
\makethesistitle

% Declaration
\begin{declaration}
    I hereby declare that this thesis is my own work, and that no other sources have been used except those indicated in the bibliography. I confirm that generative AI tools were not used in writing any part of this thesis, though they may have been used for preliminary research purposes only.

    \vspace{3cm}

    \noindent
    Place, Date \hfill Signature

    \vspace{1cm}

    \noindent
    \rule{4cm}{0.5pt} \hfill \rule{6cm}{0.5pt}
\end{declaration}

% Abstract
\begin{thesisabstract}
    % This bachelor thesis explores applications of fixed point theory
    % in mathematical analysis, with particular focus on
    % the Banach fixed point theorem and its generalizations.
    % The work investigates both theoretical foundations and practical
    % applications of fixed point theorems in solving equations and
    % optimization problems.

    % The main contributions of this thesis include a comprehensive
    % survey of classical fixed point theorems, detailed proofs of key
    % results, and several novel applications to problems in numerical
    % analysis and differential equations. We demonstrate how fixed
    % point theory provides a unifying framework for understanding
    % convergence properties of iterative methods and existence theorems
    % for solutions of functional equations.

    % The thesis concludes with a discussion of recent developments in the
    % field and potential directions for future research, including extensions
    % to non-metric spaces and applications to machine learning algorithms.

    This thesis serves as an introduction to formal mathematics using the
    Lean proof assistant. After a brief overview of the Curry--Howard
    correspondence, we explore how mathematical structures and
    properties are defined and utilized in Lean and Mathlib.
    Finally, we present a formalization of the topologist's sine curve,
    which has been merged into the Mathlib library,
    contributing to the broader Lean community.

    During the work on this thesis, I have explored new concepts
    in both mathematics and computer science, starting from logic and
    type theory within the Curry--Howard correspondence.
    The practice of formal and constructive mathematics has influenced
    my approach to mathematical reasoning. In constructing the rational
    numbers in Lean, I explored the algebraic construction using quotients,
    which is reflected in Lean through quotient types. I further examined
    how algebraic structures such as groups and rings are defined using
    structures and type classes. For more technical depth, I studied
    filters to generalize the notion of convergence in topology.
    The formalization of the topologist's sine curve has deepened
    my understanding of connectedness, closures, and continuity in topology.


\end{thesisabstract}

% Acknowledgments
\begin{acknowledgments}
    I would like to express my sincere gratitude to Prof. David Loeffler for
    the consistent support.
    His mentorship has been particualary beneficial in enlightening my path through
    formal mathematics, shaping the thesis and giving me a chance to
    actually contribute to Lean community.

    Moreover, I deeply thank FernUni for the opportunity to pursue my
    studies. I feel honored to have been able to study
    mathematics at this stage of my life through a flexible and
    high-quality program.

    The expertise of all professors have
    been exceptional, and I hope to continue my studies in the future.
\end{acknowledgments}

% Table of contents
\tableofcontents

\input{intro}
\input{logic_prop_as_types}
\input{describe_use_properties}
\input{topologist_sine_curve}

\printbibliography
\end{document}